\documentclass[journal, onecolumn]{IEEEtran}

\usepackage{cite}
\usepackage{authblk}
\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{url}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

% paper title
\title{{\Large \textbf{AI in Fintech: Portfolio Formation and Trend Prediction}}}

% author names and affiliations
\author[1]{Qingsen Zhang}
\author[1]{Baiyi Zhang}

% For affiliations, directly adjust the font size using \small, \footnotesize, etc.
\affil[1]{{\small \{zqs, baiyi\}@vt.edu}}
\affil[1]{{\small Department of Computer Science, Virginia Polytechnic Institute and State University, Falls Church, VA, USA}}

% make the title area
\maketitle

% % abstract
% \begin{abstract}
% These instructions give you guidelines for preparing papers for IEEE Transactions and Journals. 
% \end{abstract}

% % Note that keywords are not normally used for peerreview papers.
% \begin{IEEEkeywords}
% IEEE, journal, \LaTeX, paper, template.
% \end{IEEEkeywords}

\section{Problem Statement}
The advent of cryptocurrencies has revolutionized the financial landscape, offering a new asset class that is characterized by high volatility and potential for significant returns. With over 2000 crypto assets available, investors are faced with the daunting task of portfolio selection and trend prediction, both of which are critical for successful investment strategies.

Our project aims to address these challenges by leveraging artificial intelligence (AI) in two key tasks:
\begin{itemize}
  \item Portfolio Selection: Given a dataset containing the daily price of over 2000 crypto assets from the past 11 years, the first task is to select a subset of these assets to form a portfolio. 
  \item Trend Prediction: Once a portfolio is formed, the next challenge is to predict its future trend. This is a complex task due to the inherent volatility and unpredictability of crypto markets.
\end{itemize}

Our approach leverages AI in both horizontal and vertical interactions of the assets to maximize the power of AI in investment. Horizontally, AI is used across different assets to identify the optimal combination for the portfolio. Vertically, AI is used to analyze the time-series data of each asset and predict its future trend.

\section{Methodology}
\subsection{Data Retrieval}
In the process of data acquisition, we utilize the API provided by Crypto.io, a free alternative to the commonly used, but paid CoinMarketCap.com API. Our first step involves downloading a comprehensive list of all crypto assets available on Crypto.io. Subsequently, we retrieve data for each asset, which includes daily price, opening and closing values, adjusted closing values, and volume. This data undergoes further processing, where we compute the daily return, momentum, and the 52-week high ratio, among other metrics, for subsequent utilization. This approach allows us to leverage the wealth of information available in the crypto market to inform our AI-driven investment strategies.

\subsection{Portfolio Construction}
This is a combinatorial optimization problem \cite{lim2022dynamic, bartram2020artificial, chan2002artificial, gunjan2023brief}, which we propose to solve using Constraint Satisfaction Problem (CSP) techniques. CSP techniques are well-suited to problems where a solution requires satisfying a number of constraints. The goal is to find an optimal combination of assets that maximizes expected returns while minimizing risk, subject to various constraints such as budget limits and diversification requirements.

\subsection{Trend Prediction}
This is a time-series forecasting problem, which can be approached using various AI techniques. These include reinforcement learning methods like Q-search, traditional machine learning algorithms such as Multi-Layer Perceptron (MLP), and deep learning models like Long Short-Term Memory (LSTM) \cite{10007138, 10356083, oshingbesan2022modelfree, zhang2022deep, chen2023deep, hansun2022multivariate, gupta2024forecasting}. The goal is to accurately predict future price trends based on historical data, which can inform investment decisions and risk management strategies.

\subsection{Performance Evaluation}
The performance of our AI models will be assessed using a variety of financial metrics, including the Sharpe Ratio, Treynor Ratio, Jensen's Alpha, and Maximum Drawdown. These indices are integral to our two-pronged approach in the volatile cryptocurrency market. They offer a comprehensive view of the risk-return trade-off, helping us to understand the performance of our models in different market conditions.

\section{Experiments}
\subsection{Overview}
The purpose of this section is to employ the LSTM model to evaluate the performance of various portfolios selected in the previous section. Our dataset features columns for date, price, and other calculated values over a specific period. The target variable, return rate, is based on logarithmic calculations to prevent distortion in the model output. The dataset includes date-specific features that encapsulate information from a designated time frame. 

For instance, let $d_{n}$ and $r_{n}$ represent the $n$th day from the start date of the dataset and the log based return rate of $d_{n}$ respectively, the column \texttt{sum\_return\_7d} for date  represents $\sum_{i=n-7}^{n-1} r_{i} $; the \texttt{mean\_return\_7d} column calculates the average return over the same period, mathematically representing as $\frac{\sum_{i=n-7}^{n-1} r_{i} }{7}$. Additionally, the \texttt{min} and \texttt{max} columns capture the minimum and maximum return rates within this seven-day period. The column \texttt{value\_7d\_ago} refers to the value from seven days prior to $d_{n}$.

We are interested in time frames of 7, 30, 180, and 365 days. For each candidate portfolio, we construct a corresponding dataset named as \texttt{portfolio\_$n$d} to serve as our training dataset. For example, the \texttt{portfolio\_7d} dataset includes the log-based return rate for each day, along with the sum, average, minimum, and maximum of the 7-day period leading up to that day, as well as the log-based return rate from 7 days prior.

After constructing the training dataset, we split it in half. Due to the sequential nature of LSTM models, we cannot simply divide the data and apply k-fold methods to tune model parameters. Instead, we devised a 50:50 split approach: using the first half of the time series to train the model and the second half to test it. This method ensures that the sequence integrity is maintained, and it allows us to use a substantial portion of the data for evaluating the result of the portfolio.

We implemented the LSTM model using Keras. For each asset or portfolio, we started with Keras' Sequential model, to which we added an LSTM layer with 50 memory units (neurons). We set \texttt{return\_sequences} to True to ensure that the output sequence is passed to the next layer. To prevent overfitting, we introduced a random dropout rate of 0.2 in each layer, meaning that 20\% of the neurons are dropped during training. We incorporated a total of three LSTM layers, as we consider cryptocurrency price prediction to be a complex problem. For optimization, we selected Adam, and we used mean squared error as the loss function. The entire training dataset will be trained over 100 epochs.

% references section
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,tex/proposal/ref}

\end{document}